---
title: "Data Science Final"
format: pdf
---
```{r, echo = FALSE}
library("RSocrata")
library(tidyverse)
library(stringr)
library(lubridate)
library(sf)
library(dplyr)
library(ggplot2)
library(patchwork)
```



## Water point data
We download the data using a KPI. We then filter to keep data on Cox's Bazar exclusively.
```{r}
#Downloading the data 
df <- read.socrata(
  "https://data.waterpointdata.org/resource/eqje-vguj.json?country_id=BGD",
  app_token = "your token",
  email     = "email you signed up with",
  password  = "your password"
)

#Filtering the data
filtered_df <- df %>%
  filter(!is.na(lat_deg), !is.na(lon_deg)) %>%
  filter(clean_adm2 == "Cox'S Bazar")


filtered_df <- st_as_sf(filtered_df, coords = c("lon_deg", "lat_deg"), remove = FALSE) %>%
  st_set_crs(value = 4326)

#Ploting the map:
ggplot() +
  geom_sf(data = filtered_df, aes(color = status_clean))
```


## Administrative division data
https://data.humdata.org/dataset/401d3fae-4262-48c9-891f-461fd776d49b/resource/08736818-ae72-44a9-abd6-a51915c24921/download/bgd_adm_bbs_20201113_shp.zip

This database gives us geospatial information of Bangladesh's administrative divisions.
```{r}
# Load the districts shapefile
districts <- st_read("path to your download")

# Filter the data
filtered_districts <- districts %>%
  filter(ADM2_EN == "Cox's Bazar")

#Plot the map:
ggplot() +
  geom_sf(data = filtered_districts, aes(color = ADM3_EN))

```

## Downloading refuge camp data
https://data.humdata.org/dataset/401d3fae-4262-48c9-891f-461fd776d49b/resource/08736818-ae72-44a9-abd6-a51915c24921/download/bgd_adm_bbs_20201113_shp.zip

This data set gives us geographical information on refugee camps in Cox's Bazar, at different levels of detail.
```{r}
camp_data <- st_read("path to your download")

camp_layers <- st_layers(dsn = "path to your download")
camp_layers

camp_outlines <- st_read("path to your download", "T20230412_A1_Camp_Outlines")
plot(camp_outlines)

camp_block_outlines <- st_read("path to your download", "T20230412_A2_Camp_Block_Outlines")
plot(camp_block_outlines)

camp_sub_block_outlines <- st_read("path to your download", "T20230412_A3_Camp_SubBlock_Outlines")
plot(camp_sub_block_outlines)
```

## Combining the maps

```{r}
# Combining the water point and district divisions maps
ggplot() +
  geom_sf(data = filtered_districts, aes(fill = ADM3_EN), color = "black") +
  scale_fill_discrete() +
  
  geom_sf(data = filtered_df, aes(color = status_clean)) +
  scale_color_manual(values = c("Functional" = "green", "Non-Functional, dry season" = "blue", "Non-Functional" = "orange")) +

  labs(fill = "Districts Name", color = "Water point Status")

```


```{r}
# Combining the refuge camps and district maps
ggplot() +
  geom_sf(data = filtered_districts, aes(fill = ADM3_EN), color = "black") +
  geom_sf(data = camp_outlines, fill = "transparent", color = "yellow") +
  scale_fill_discrete(name = "Districts Name") +
  scale_color_manual(name = "Refugee camp outlines", values = "yellow")

```

Note: mapping all 3 things together is not possible because of overlapping data (water points and refugee camps)

## Function: maps for different districts
```{r}

plot_filtered_district <- function(district_filter, layer, water_point_filter, camp_outline_filter, Title, Level) {
  # Apply filters to the district and water point data
  new_filtered_districts <- dplyr::filter(filtered_districts, district_filter)
  new_filtered_df <- dplyr::filter(filtered_df, water_point_filter)
  new_filtered_camp_outlines <- dplyr::filter(camp_outlines, camp_outline_filter)
  
  # Plotting
  ggplot() +
    geom_sf(data = new_filtered_districts, aes(fill = !!as.name(layer)), color = "black") +
    geom_sf(data = new_filtered_df, aes(color = status_clean)) +
    scale_color_manual(values = c("Functional" = "green", "Non-Functional, dry season" = "blue", "Non-Functional" = "orange")) +
    geom_sf(data = new_filtered_camp_outlines, fill = "transparent", color = "yellow") +
    scale_fill_discrete(name = Level) +
    labs(
      color = "Water point Status",
      title = Title) +
    theme_minimal()
  
}


```

### Plots for different districts
```{r}
Ukhia <- plot_filtered_district(district_filter = filtered_districts$ADM3_EN == "Ukhia",
                                layer = "ADM3_EN",
                                water_point_filter = filtered_df$clean_adm3 == "Ukhia",
                                camp_outline_filter = camp_outlines$Upazila == "Ukhia",
                                "Ukhia District",
                                "District Name")


Teknarf <- plot_filtered_district(district_filter = filtered_districts$ADM3_EN == "Teknaf",
                                  layer = "ADM3_EN",
                                  water_point_filter = filtered_df$clean_adm3 == "Teknaf",
                                  camp_outline_filter = camp_outlines$Upazila == "Teknaf",
                                  "Teknaf District",
                                  "District Name")
Ukhia + Teknarf
```

### Plots for different districts including sub district breakdown
```{r}
Ukhia <- plot_filtered_district(district_filter = filtered_districts$ADM3_EN == "Ukhia",
                                layer = "ADM4_EN",
                                water_point_filter = filtered_df$clean_adm3 == "Ukhia",
                                camp_outline_filter = camp_outlines$Upazila == "Ukhia",
                                "Ukhia District",
                                "Sub-district name")

Teknarf <- plot_filtered_district(district_filter = filtered_districts$ADM3_EN == "Teknaf",
                                  layer = "ADM4_EN",
                                  water_point_filter = filtered_df$clean_adm3 == "Teknaf",
                                  camp_outline_filter = camp_outlines$Upazila == "Teknaf",
                                  "Teknaf District",
                                  "Sub-district name")

Ukhia + Teknarf
```

We now need to dissagregate further, since there is still a lot of overlap between water points and refuge camps data. The sub districts I would prioritize visualizations for would be:

In Ukhia: Palong Khali

In Teknaf:  Whykong, Nhilla, Teknaf

